---
title: "236523 - Introduction to Bioinformatics - Spring 21 - HW2"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Roy Bernea - 316315829 ; Alon Hacohen - 311587653

# Qusetion 1

# TODO
Following is the code and answers for the questions under the "Question 1" Section - as instructed in the assignment:

### Q1.1
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

```

# Question 2

# TODO - Roy
Following is the code and answers for the questions under the "Question 2" Section - as instructed in the assignment:

### Q2.1
# TODO


### Q2.2
# TODO

### Q2.3
### Q2.3.1
Additional names (aliases) for the gene *MARCH7* Are:
Membrane Associated Ring-CH-Type Finger
Axotrophin
MARCH-VII
RNF177
Membrane-Associated Ring Finger (C3HC4) 7, E3 Ubiquitin Protein Ligase
RING-Type E3 Ubiquitin Transferase MARCHF7
Membrane-Associated RING Finger Protein 7
Membrane-Associated RING-CH Protein VII
E3 Ubiquitin-Protein Ligase MARCHF7
RING Finger Protein 177
MARCH7
AXOT
Membrane-Associated Ring Finger (C3HC4)
RING-Type E3 Ubiquitin Transferase MARCH7
E3 Ubiquitin-Protein Ligase MARCH7
Membrane Associated Ring Finger 7
EC 2.3.2.27
EC 6.3.2 
MARCHF7
AXO

These gene names were extracted from the following databases:
#TODO

### Q2.3.2
If we paste the name of this gene manually into an Excel sheet, meaning we paste "MARCH7" into a cell inside the sheet, a possible problem we might encounter is that Excel would recognize the name as a date - March 7 - and refer to it as one.
In such case, instead of referring to the gene's name "MARCH7" as its ID (or as a unique identifier), Excel would mistakenly refer to it as a date - which might cause us problem later on in our data analysis process (e.g. if we want to compare the gene in our excel sheet to its appearance in a database - but R would recognize the identifiers as two different ones because of their type etc.). 




# Question 3

# TODO
Following is the code for generating the GRanges objects - as instructed in the assignment:

### Q3.1
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

library(ggplot2)
library(RColorBrewer)
library(factoextra)
library(tidyverse)
library(Rtsne)
library(dbscan)

# data preperation
#change to relative path?
sdy <- readRDS("C:/CS_Technion/236523-Intro_to_Bioinformatics/Spr21/Intro_to_BioInf-Spr21-HW2/sdy420.rds")
#print(t(sdy$expr))
data <- na.omit(t(sdy$expr))
data.scaled <- scale(data)
#view(data)
#dist.mat <- dist(data)

# preparing tsne for later
tsne = Rtsne(data.scaled)

# Clustering methods

# kmeans
km.result <- kmeans(data.scaled, centers = 3, nstart = 25)
# hierarchical clustering
hc.result <- hclust(dist(data.scaled), "ward.D")
# DBSCAN
dbscan.result <- dbscan(tsne$Y, eps = 1, minPts = 4)


# Plotting the clusters

#plot(tsne$Y, col=km.result$cluster)
#plot(tsne$Y, col=hc.result$order)
#plot(tsne$Y, col=dbscan.result$cluster)

#dbscan.result


```


order
cluster